---
author: ""
date: ""
title: "Analysis of Sport Team Colors, Part 2"
output:
  html_document:
    toc: false
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  echo = TRUE,
  # echo = FALSE,
  # cache = TRUE,
  # results = "markdown",
  # results = "hide",
  # fig.align = "center",
  # fig.show = "asis",
  fig.width = 10,
  fig.height = 6,
  # out.width = 6,
  # out.height = 6,
  warning = FALSE,
  message = FALSE
)

```

This write-up picks up where the previous one left off. All of the session data
is carried over.

```{r restore_session}
# session::restore.session(file.path("data", "colors-eda-1.RData"))
load(file.path("data", "colors-eda-1.RData"))
# vars_all <- ls()
# vars_keep <- stringr::str_subset(vars_all, "^add_|^convert_|^lab_|^viz_*_base$|(colors_tidy_ord2)")
# rm(list = setdiff(vars_all, vars_keep))
library("dplyr")
library("ggplot2")
```

## Color Similarity

Now, I'd like to evaluate color similarity more closely. To help verify 
any quantitative deductions with some intuition,
I'll consider only a single league for this--the NBA, the league that I know the best. 

Because I'll end up plotting team names at some point (probably with a heat map), I want
to get the official abbreviations for each team (because the full names are relatively lengthy).
Unfortunately, these
don't come with the [`teamcolor` package](https://github.com/beanumber/teamcolors). 
Nevertheles, I can use [Alex Bresler](http://asbcllc.com/)'s 
[`nbastatR` package](https://github.com/abresler/nbastatR) to do so. 

```{r nms_nba}
# Assign `df_dict_nba_teams` to Global environment.
nbastatR::assign_nba_teams()
nms_nba <-
  teamcolors::teamcolors %>% 
  filter(league == "nba") %>% 
  inner_join(
    df_dict_nba_teams %>%
      setNames(snakecase::to_snake_case(names(.))) %>%
      filter(!is_non_nba_team) %>% 
      select(name = name_team, slug = slug_team),
    by = c("name")
  )

colors_tidy_ord2_nba <-
  nms_nba %>% 
  select(name, league, slug) %>% 
  inner_join(colors_tidy_ord2, by = c("name", "league"))
```

To give the unfamiliar reader a better understanding of what exactly this subset of the
`teamcolors` data incorporate, here's a 
visualization of the primary and secondary colors of all NBA teams

```{r viz_colors_ord2_nba, echo = FALSE, fig.heigh = 10, fig.width = 6}
colors_ord2_nba <-
  colors_tidy_ord2_nba %>% 
  tidyr::spread(ord, hex)

lab_title_colors_ord2_nba <-
  stringr::str_replace_all(lab_title_colors_all, lab_lgs, "NBA")

# Reference:
# + https://github.com/beanumber/teamcolors.
viz_colors_ord2_nba <-
  colors_ord2_nba  %>% 
  mutate(name_slug = paste0(name, ": ", slug)) %>% 
  ggplot(aes(x = name_slug, color = name_slug, fill = name_slug)) + 
  geom_bar(size = 1.5) +
  coord_flip() +
  scale_x_discrete(NULL) + 
  scale_y_continuous(NULL) + 
  scale_fill_manual(values = colors_ord2_nba$primary) + 
  scale_color_manual(values = colors_ord2_nba$secondary) + 
  viz_theme_base +
  theme(axis.text.x = element_blank(), 
        axis.ticks.x = element_blank()) +
  viz_labs_base +
  labs(title = lab_title_colors_ord2_nba)
viz_colors_ord2_nba
```


After grabbing the abbreviations (or `slug`s), I can move on to breaking up
the hex values into their RGB components. (I use the same method as the one used before.)
I'll be looking at only the primary and secondary colors again.

```{r colors_ord2_nba_rgb_tidy}
colors_ord2_nba_rgb_tidy <-
  colors_tidy_ord2_nba %>%
  add_rgb_cols() %>% 
  select(-hex) %>%
  tidyr::gather(rgb, value, red, green, blue)

colors_ord2_nba_rgb_tidy %>% 
  create_kable()
```

```{r ggimage, include = FALSE, eval = FALSE}
# Reference:
# + https://cran.r-project.org/web/packages/ggimage/vignettes/ggimage.html.
n <- 3
path_img <- file.path("data-raw", "sas.png")
path_img <- "http://i.cdn.turner.com/nba/nba/.element/media/2.0/teamsites/spurs/Schad/img/PostgameTeams/SAS.png"
path_img_dwnld <- "https://github.com/jimniels/teamcolors/blob/master/static/img/nfl/arizona-cardinals.svg"
convertGraph::convertGraph(path_img_dwnld, "out.png")
# path_img <- "content.sportslogos.net/logos/7/177/full/7420_arizona_cardinals-primary_on_dark-2005.png"
set.seed(42)
d <- data.frame(
  x = rnorm(n),
  y = rnorm(n),
  img = path_img, 
  size = n
)
d %>% 
  ggplot(aes(x, y)) + 
  ggimage::geom_image(aes(image = img), size = .05)

```

With the RGB values extracted, I can use the `widyr::pairwise_dist()` function
to compute the relative distance among teams in terms of RGB values for each color 
ordinality. [^fn_pairwise_dist] [^fn_pairwise_similarity]
I think the default method--["Euclidean" distance](https://en.wikipedia.org/wiki/Euclidean_distance)--is reasonable.
But, because I'm curious, I'll look at how different the results would be if the
["Manhattan" distance](https://en.wiktionary.org/wiki/Manhattan_distance) is used instead.

```{r colors_ord2_nba_rgb_dist}
do_pairwise_dist <- function(data, method) {
  data %>% 
    group_by(ord) %>% 
    widyr::pairwise_dist(name, rgb, value, upper = TRUE, method = method) %>% 
    rename(name1 = item1, name2 = item2) %>% 
    select(everything(), value = ncol(.)) %>% 
    arrange(value, .by_group = TRUE) %>% 
    ungroup()
}

correlate_pairwise_out2 <- function(data1, data2) {
  corrr::correlate(
    data1 %>% 
      rename(value1 = value) %>% 
      left_join(data2 %>% rename(value2 = value)) %>%
      select_if(is.numeric)
  )
}

correlate_pairwise_out2(
  colors_ord2_nba_rgb_tidy %>%
    do_pairwise_dist("euclidean"),
  colors_ord2_nba_rgb_tidy %>%
    do_pairwise_dist("manhattan") 
) %>% 
  create_kable()
```

As one might expect, there's not much difference between these 
two distance methods (if correlation is deemed a valid metric for similarity).
Some other distance methods would probably result in a more dissimilar result set,
but these two methods seem most appropriate.

How exactly do all of the individual distances compare?

```{r viz_colors_ord2_nba_rgb_dist}
colors_ord2_nba_rgb_dist <-
  colors_ord2_nba_rgb_tidy %>%
    do_pairwise_dist("euclidean")

colors_ord2_nba_rgb_dist_ranks <-
  colors_ord2_nba_rgb_dist %>%
  group_by(ord, name1) %>%
  summarize(value = mean(value)) %>%
  ungroup() %>%
  group_by(name1) %>% 
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  mutate(rank = row_number(value)) %>% 
  arrange(rank)
```

```{r viz_colors_ord2_nba_rgb_dist}

# add_lg_and_slug_cols <- function(data) {
#   data %>%
#   left_join(nms_nba %>% select(name1 = name, slug1 = slug, league1 = league), by = c("name1")) %>%
#   left_join(nms_nba %>% select(name2 = name, slug2 = slug, league2 = league), by = c("name2"))
# }

colors_ord2_nba_rgb_dist_ranks

lvls_ord2_nba_rgb_dist_ranks <-
  colors_ord2_nba_rgb_dist_ranks %>%
  pull(name1)

labs_ord2_nba_rgb_dist_ranks <-
  colors_ord2_nba_rgb_dist_ranks %>%
  left_join(nms_nba %>% select(name1 = name, slug1 = slug, league1 = league), by = c("name1"))
  pull(slug1)

viz_colors_ord2_nba_rgb_dist <-
  colors_ord2_nba_rgb_dist %>% 
  mutate(name1 = factor(name1, levels = lvls_ord2_nba_rgb_dist_ranks)) %>% 
  ggplot(aes(x = name1, y = value)) +
  geom_point(alpha = 0.25, color = "grey50") +
  geom_point(
    data = colors_ord2_nba_rgb_dist_ranks,
    aes(color = name1),
    # color = "blue",
    size = 10
  ) +
  scale_color_manual(values = colors_ord2_nba$primary) +
  scale_x_discrete(labels = labs_ord2_nba_rgb_dist_ranks) +
  coord_polar(theta = "x") +
  viz_theme_base +
  theme(axis.ticks = element_blank(), axis.text.y = element_blank()) +
  viz_labs_base
viz_colors_ord2_nba_rgb_dist
```

```{r colors_ord2_nba_rgb_dist}
colors_ord2_nba_rgb_dist <-
  colors_ord2_nba_rgb_tidy %>% 
  do_pairwise_dist("euclidean")
```

```{r colors_ord2_nba_rgb_dist_debug, include = FALSE, eval = FALSE}
colors_ord2_nba_rgb_dist %>% 
  group_by(ord, name1) %>% 
  do(slice(., c(1, n()))) %>% 
  ungroup()
colors_ord2_nba_rgb_dist %>% 
  arrange(value) %>%
  filter(ord == "primary") %>% 
  filter(value != 0) %>% 
  slice(c(1:10, c(n()-9):n())) %>% 
  ungroup()
colors_ord2_nba_rgb_tidy %>%
  filter(name %in% c("San Antonio Spurs", "Pittsburgh Steelers", "Brooklyn Nets")) %>%
  tidyr::spread(rgb, value)
```

I can use a heatmap to visualize the distance values as well.
In order to make the plot more insightful, I order the teams on each axis
by total distance from all other teams--teams with the highest cumulative similarity to all
other teams appear towards the bottom and left, while teams that
contrast most with all others appear towards the top and right. Additionally,
to add some nuance, I emphasize the individual team-pairs that have the highest
and lowest similarity.


```{r viz_colors_ord2_nba_rgb_dist_byord_setup, echo = FALSE}
add_lg_and_slug_cols <- function(data) {
  data %>%
  left_join(nms_nba %>% select(name1 = name, slug1 = slug, league1 = league), by = c("name1")) %>%
  left_join(nms_nba %>% select(name2 = name, slug2 = slug, league2 = league), by = c("name2"))
}

colors_ord2_nba_rgb_dist_ranks <-
  colors_ord2_nba_rgb_dist %>%
  add_lg_and_slug_cols() %>% 
  group_by(ord, name1, slug1) %>%
  summarize(value = mean(value)) %>%
  ungroup() %>%
  group_by(name1, slug1) %>% 
  summarize(value = mean(value)) %>% 
  ungroup() %>% 
  mutate(rank = row_number(value)) %>% 
  arrange(rank)

lvls_ord2_nba_h2h_ranks <-
  colors_ord2_nba_rgb_dist_ranks %>%
  pull(name1)

labs_ord2_nba_h2h_ranks <-
  colors_ord2_nba_rgb_dist_ranks %>%
  pull(slug1)
```

```{r colors_ord2_nba_rgb_dist_byord_viz, echo = FALSE}
create_colors_ord2_nba_rgb_dist_viz <- function(data, ranks) {
  data %>% 
    # tidyr::complete(name1, name2) %>% 
    inner_join(
      ranks %>% select(name1, rank1 = rank),
      by = c("name1")
    ) %>% 
    inner_join(
      ranks %>% select(name2 = name1, rank2 = rank),
      by = c("name2")
    ) %>% 
    # mutate_at(vars(name1, name2), funs(factor(., levels = lvls_ord2_nba_h2h_ranks)))
    mutate(name1 = reorder(name1, rank1),
           name2 = reorder(name2, rank2))
}
colors_ord2_nba_rgb_dist_byord_viz <-
  colors_ord2_nba_rgb_dist  %>%
  create_colors_ord2_nba_rgb_dist_viz(colors_ord2_nba_rgb_dist_ranks)
```


```{r colors_ord2_nba_rgb_dist_byord_viz_debug, include = FALSE, eval = FALSE}
# colors_ord2_nba_rgb_dist_viz %>% arrange(rank1)
# colors_ord2_nba_rgb_dist_viz %>% arrange(name1)
# levels(colors_ord2_nba_rgb_dist_viz$name1)
# levels(colors_ord2_nba_rgb_dist_viz$name2)
```

```{r viz_colors_ord2_nba_rgb_dist_setup, echo = FALSE}
lab_caption_ord2_nba_h2h <-
  paste0(
    "Dark = similar; light = dissimilar.",
    "Similarity is based on Euclidean distance of RGB values.\n",
    " Order is determined by overall similarity across both primary and secondary colors.",
    "\n Top 2% similar and dissimilar team-team colors emphasized with black and white respectively."
  )
viz_labs_colors_ord2_nba_rgb_dist <-
  labs(
    title = "Pairwise Distance of Colors",
    subtitle = lab_subtitle_ord2_nba,
    caption = lab_caption_ord2_nba_h2h,
    x = NULL,
    y = NULL
  )

visualize_colors_ord2_nba_rgb_dist_const <-
  function(data, value = "value", emphasize_top = TRUE, pct_top = 0.02) {

    value_sym <- rlang::sym(value)
    viz <-
      data %>% 
      # filter(league1 == league2) %>%
      # mutate(lg_ord = paste0(toupper(league1), ", ", stringr::str_to_title(ord))) %>% 
      mutate(!!value_sym := ifelse(is.na(!!value_sym), 0, !!value_sym)) %>% 
      ggplot(aes(x = name1, y = name2)) +
      geom_tile(aes_string(fill = value_sym)) +
      scale_x_discrete(labels = labs_ord2_nba_h2h_ranks) +
      scale_y_discrete(labels = labs_ord2_nba_h2h_ranks) +
      viridis::scale_fill_viridis(option = "E") +
      # scale_fill_gradient(low = "black", high = "white") +
      viz_theme_base +
      theme(axis.text.x = element_text(angle = 90)) +
      viz_labs_colors_ord2_nba
    
    if(emphasize_top) {
      data_pranked <-
        data %>% 
        mutate(prank = percent_rank(!!value_sym))
      viz <-
        viz +
        geom_tile(
          data = data_pranked %>%
            filter(prank <= pct_top),
          fill = "green"
        ) +
        geom_tile(
          data = data_pranked %>%
            filter(prank >= (1 - pct_top)),
          fill = "red"
        )
    }
    viz
  }
```

```{r viz_colors_ord2_nba_rgb_dist_byord, include = FALSE, eval = FALSE}
viz_colors_ord2_nba_rgb_dist_byord <-
  colors_ord2_nba_rgb_dist_byord_viz %>%
  visualize_colors_ord2_nba_rgb_dist_const() +
  facet_wrap(~ ord, scales = "free")
viz_colors_ord2_nba_rgb_dist_byord
```


```{r colors_ord2_nba_rgb_dist_viz, include = FALSE}
colors_ord2_nba_rgb_dist_viz <-
  colors_ord2_nba_rgb_dist %>%
  group_by(name1, name2) %>% 
  summarize(value = sum(value)) %>% 
  ungroup() %>% 
  create_colors_ord2_nba_rgb_dist_viz(colors_ord2_nba_rgb_dist_ranks)
```

```{r viz_colors_ord2_nba_rgb_dist, echo = FALSE}
viz_colors_ord2_nba_rgb_dist <-
  colors_ord2_nba_rgb_dist_viz %>%
  visualize_colors_ord2_nba_rgb_dist_const()
viz_colors_ord2_nba_rgb_dist
```

Exactly which teams make up the top and bottom percent (in terms of color similarity)?
Here's a list of the top and bottom 10. [^fn_topbot_ties]

[^fn_topbot_ties]:
There are some ties--these are broken by cumulative similarity rank (i.e. the metric shown in the previous plot),
which is not shown in the data printout.

```{r colors_ord2_nba_rgb_dist_viz_prank, echo = FALSE, results = "markdown"}
colors_ord2_nba_rgb_dist_viz_prank <-
  colors_ord2_nba_rgb_dist_viz %>% 
  mutate(prank = percent_rank(value)) %>% 
  arrange(prank)

slice_topbot <- function(data, num) {
  data %>%
    slice(c(1:num, c(n() - (num - 1)):n()))
}

colors_ord2_nba_rgb_dist_viz_prank %>% 
  select(-rank1, -rank2, -prank) %>% 
  slice_topbot(10) %>% 
  create_kable()
```



But is the [RGB model](https://en.wikipedia.org/wiki/RGB_color_model) 
really the best framework to use for comparing colors? What about the 
[HSL color schema](https://en.wikipedia.org/wiki/HSL_and_HSV) (Hue, Saturation, Lightness), 
which incorporates hue and lightness (two concepts that I've looked at already)?
While the HSL model is worth evaluating, my research on this topic--which
mostly consisted of a quick Google search for
["What is the best method for identifying similarity between colors?"](https://stackoverflow.com/questions/5392061/algorithm-to-check-similarity-of-colors)--indicates
that another encoding might be best--the [YUV representation](https://en.wikipedia.org/wiki/YUV).

Since' I'm not expert in color models, I think its a good idea to perform distance
calculations using each schema and see how well they agree with one another.

```{r yuv_funcs, echo = FALSE}
convert_rgb2yuv <- function(r, g, b) {
  out <- colorscience::RGB2YUV(c(r, g, b))
  c(y = out[1], u = out[2], v = out[3])
}

convert_hex2yuv <- function(hex, unname = FALSE) {
  rgb <- t(grDevices::col2rgb(hex))
  out <- convert_rgb2yuv(rgb[1], rgb[2], rgb[3])
  if(unname) {
    out <- unname(out)
  }
  out
}

add_yuv_cols <- function(data) {
  data %>%
    pull(hex) %>%
    purrr::map(convert_hex2yuv) %>% 
    purrr::reduce(rbind) %>% 
    tibble::as_tibble() %>%
    bind_cols(data, .)
}
```

To begin, I'll create "tidy" HSL and YUV data sets 
(using some custom functions that I do not show).
Then, as a sort-of "sanity check", I'll verify that the similarity determined by
two distance methods--Euclidean and Manhattan--is nearly identical for each data set
(corresponding to a singular color model).


```{r colors_ord2_nba_hsl_yuv_dist}
colors_ord2_nba_hsl_tidy <-
  colors_tidy_ord2_nba %>%
  add_hsl_cols() %>% 
  select(-hex) %>%
  tidyr::gather(rgb, value, h, s, l)

colors_ord2_nba_yuv_tidy <-
  colors_tidy_ord2_nba %>%
  add_yuv_cols() %>% 
  select(-hex) %>%
  tidyr::gather(rgb, value, y, u, v)

correlate_pairwise_out2(
  colors_ord2_nba_hsl_tidy %>%
    do_pairwise_dist("euclidean"),
  colors_ord2_nba_hsl_tidy %>%
    do_pairwise_dist("manhattan")
) %>% 
  create_kable()

correlate_pairwise_out2(
  colors_ord2_nba_yuv_tidy %>%
    do_pairwise_dist("euclidean"),
  colors_ord2_nba_yuv_tidy %>%
    do_pairwise_dist("manhattan")
) %>% 
  create_kable()
```

Now, when keeping the distance method constant,
what if I compare the calculated values across the each underlying
data set (corresponding to the different color models)?

```{r colors_ord2_nba_rgb_hsl_yuv_dist, echo = FALSE}
cor_dist_data <-
  colors_ord2_nba_rgb_tidy %>% 
  do_pairwise_dist("euclidean") %>% 
  rename(rgb_dist = value) %>% 
  left_join(
    colors_ord2_nba_hsl_tidy %>% 
      do_pairwise_dist("euclidean") %>% 
      rename(hsl_dist = value)
  ) %>%
  left_join(
    colors_ord2_nba_yuv_tidy %>% 
      do_pairwise_dist("euclidean") %>% 
      rename(yuv_dist = value)
  )

# # A "tidy" method (i.e. not using `stats::cor()` that doesn't produce a plot.
# cor_dist_data %>% 
#   select_if(is.numeric) %>% 
#   corrr::correlate()
lab_title_cor_dist <- "Correlation Matrix for RGB, HSL, and YUV Euclidean Distance Calculations"
lab_subtitle_ord2_nba <-
  stringr::str_replace_all(lab_subtitle_kclusts, lab_lgs, "NBA")
viz_cor_dist <-
  cor_dist_data %>% 
  select_if(is.numeric) %>% 
  GGally::ggcorr(label = TRUE, label_round = 2) +
  viz_theme_base +
  theme(legend.position = "right") +
  viz_labs_base +
  labs(title = lab_title_cor_dist, subtitle = lab_subtitle_ord2_nba)
viz_cor_dist
```


Even if one can decide upon a "best" color scheme to use for comparing colors,
there is the question of whether "distance" is the best method for doing so.
Why not a "similarity" metric (such as cosine similarity)?

Again, since I don't really have a strong opinion on which metric is appropriate,
I think it's worthwhile to look at both. I'll save the reader from the details--it's
not too dissimilar (pun intended) from the methods shown before--
and skip right to a visualization showing the correlations among the distance metric (Euclidean method)
and similarity metric for each color model--RGB, HSL, and YUV. (This presents
a good opportunity to use [Dr. Simon](https://drsimonj.svbtle.com/)'s 
[`{corrr} package`](https://github.com/drsimonj/corrr), including its handy `network_plot()` function.
(`GGally::ggcorr()` was used for the previous correlation plot).

```{r colors_ord2_nba_rgb_sim, include = FALSE}
do_pairwise_sim_byord <- function(data, ord) {
  data %>%
    filter(ord == ord) %>% 
    select(-ord) %>% 
    widyr::pairwise_similarity(name, rgb, value) %>% 
    rename(name1 = item1, name2 = item2) %>% 
    select(everything(), value = ncol(.)) %>% 
    mutate(ord = ord) %>% 
    select(ord, everything()) %>% 
    arrange(desc(value))
}

do_pairwise_sim <-
  function(data, ords = c("primary", "secondary")) {
    bind_rows(
      data %>%
        do_pairwise_sim_byord(ords[1]),
      data %>%
        do_pairwise_sim_byord(ords[2])
    )
  }
```

```{r corrr_rgb_hsl_dist_sim, echo = FALSE}
cor_sim_data <-
  colors_ord2_nba_rgb_tidy %>% 
  do_pairwise_sim() %>% 
  rename(rgb_sim = value) %>% 
  left_join(
    colors_ord2_nba_hsl_tidy %>% 
      do_pairwise_sim() %>% 
      rename(hsl_sim = value)
  ) %>% 
  left_join(
    colors_ord2_nba_yuv_tidy %>% 
      do_pairwise_sim() %>% 
      rename(yuv_sim = value)
  )
      
cor_distsim_data <-
  cor_dist_data %>%  
  left_join(cor_sim_data)
cor_rgbhslyuv_distsim <- 
  cor_distsim_data %>% 
  select_if(is.numeric) %>% 
  corrr::correlate()

lab_title_cor_rgbhslyuv_distsim <-
  paste0("Correlation Network for RGB, HSL, and YUV Euclidean Distance",
         "and Cosine Similarity Calculations")
viz_cor_rgbhslyuv_distsim <-
  cor_rgbhslyuv_distsim %>% 
  corrr::network_plot(
    min_cor = 0, 
    colors = c("red", "green")
  ) +
  viz_theme_base +
  theme(legend.position = "bottom", axis.ticks = element_blank(), axis.text = element_blank()) +
  viz_labs_base +
  labs(title = lab_title_cor_rgbhslyuv_distsim, subtitle = lab_subtitle_ord2_nba)
viz_cor_rgbhslyuv_distsim
```

It's clear that the RGB and YUV are fairly similar according to both metrics--Euclidean distance
and cosine similarity--and both are relatively dissimilar to HSL. However, all three color models
show negative correlations with themselves when comparing the two metrics.

___So, which color model and which metric should be used?___
In my (uninformed) opinion, the RGB model seems like a good choice, both because it is
relatively similar to at least one other method (YUV) and because it is (probably) the most
relatable scheme to people who don't know much about color theory. For metric, I think that
the choice of Euclidean distance is very defensible. My Google search (which makes the case for YUV) makes the assumption
that Euclidean distance is being used. Additionally, a second Google search
for "euclidean distance vs. cosine similarity" turns up 
[an easy-to-follow technical write-up](https://cmry.github.io/notes/euclidean-v-cosine)
that implies
that cosine similarity is probably not really appropriate for this kind of color analysis.

### Checking the Conclusions

[Maksim Horowitz](https://github.com/maksimhorowitz)'s
[`nflscrapR` package](https://github.com/maksimhorowitz/nflscrapR) to do so.


```{r nms_nfl}
# Need an alternative method of joining since the names are not
# identical across `teamcolors::teamcolors` and `nflscrapR::nflteams`.
identify_tm_mascot <- function(x) {
  stringr::str_replace_all(x, "^.*\\s", "")
}

nms_nfl <-
  teamcolors::teamcolors %>% 
  filter(league == "nfl") %>% 
  mutate(mascot = identify_tm_mascot(name)) %>% 
  inner_join(
    nflscrapR::nflteams %>%
      setNames(snakecase::to_snake_case(names(.))) %>%
      mutate_all(as.character) %>% 
      mutate(mascot = identify_tm_mascot(team_name)) %>% 
      select(slug = abbr, mascot),
    by = c("mascot")
  ) %>% 
  select(-mascot)
```
