---
author: ""
date: ""
title: "Analysis of Sport Team Colors, Part I"
output:
  html_document:
    toc: false
---


## Color Similarity

Now, I'll take a closer look at the two leagues that I follow most closely--the NBA
and the NFL. Because I'll end up plotting team names at some point, I want
to get the official abbreviations for each team. I can use data/methods
from [Alex Bresler](http://asbcllc.com/)'s 
[{nbastatR} package](https://github.com/abresler/nbastatR) and 
[Maksim Horowitz](https://github.com/maksimhorowitz)'s
[{nflscrapR} package](https://github.com/maksimhorowitz/nflscrapR) to do so.

```{r nms_nbanfl}
nbastatR::assign_nba_teams()
nms_nba <-
  teamcolors::teamcolors %>% 
  filter(league == "nba") %>% 
  inner_join(
    df_dict_nba_teams %>%
      setNames(snakecase::to_snake_case(names(.))) %>%
      filter(!is_non_nba_team) %>% 
      select(name = name_team, slug = slug_team),
    by = c("name")
  )

identify_tm_mascot <- function(x) {
  stringr::str_replace_all(x, "^.*\\s", "")
}

# Need an alternative method of joining since the names are not
# identical across `teamcolors::teamcolors` and `nflscrapR::nflteams`.
nms_nfl <-
  teamcolors::teamcolors %>% 
  filter(league == "nfl") %>% 
  mutate(mascot = identify_tm_mascot(name)) %>% 
  inner_join(
    nflscrapR::nflteams %>%
      setNames(snakecase::to_snake_case(names(.))) %>%
      mutate_all(as.character) %>% 
      mutate(mascot = identify_tm_mascot(team_name)) %>% 
      select(slug = abbr, mascot),
    by = c("mascot")
  ) %>% 
  select(-mascot)

nms_nbanfl <- bind_rows(nms_nba, nms_nfl)
```

After grabbing the abbreviations (or `slug`s), I can move on to breaking up
the hex values into their rgb components. (I use the same method as the one used before.)
I'll be looking at only the primary and secondary colors again.


```{r colors_nbanfl_ord2_rgb_tidy}
colors_nbanfl_tidy_ord2 <-
  nms_nbanfl %>% 
  select(name, league, slug) %>% 
  inner_join(colors_tidy_ord2, by = c("name", "league"))

colors_nbanfl_ord2_rgb_tidy <-
  colors_nbanfl_tidy_ord2 %>%
  add_rgb_cols() %>% 
  select(-hex) %>%
  tidyr::gather(rgb, value, red, green, blue)
colors_nbanfl_ord2_rgb_tidy
```

With the rgb values extracted, I can use the `widyr::pairwise_dist()` function
(which defaults to a "Euclidean" metric of distance) to compute the relative
distance among teams in terms of rgb values for each color ordinality. [^fn_pairwise_dist] [^fn_pairwise_similarity]

[^fn_pairwise_dist]:
I tried a "Manhattan" method and got nearly identical results.

[^fn_pairwise_similarity]:
The `widyr::pairwise_similarity()` function might also be used here.

```{r colors_nbanfl_ord2_rgb_dist}
do_pairwise_dist <- function(data, method) {
  data %>% 
    group_by(ord) %>% 
    widyr::pairwise_dist(name, rgb, value, upper = TRUE, method = method) %>% 
    rename(name1 = item1, name2 = item2) %>% 
    select(everything(), value = ncol(.)) %>% 
    arrange(value, .by_group = TRUE) %>% 
    ungroup()
}

colors_nbanfl_ord2_rgb_dist <-
  colors_nbanfl_ord2_rgb_tidy %>%
  do_pairwise_dist("euclidean")
```

```{r colors_nbanfl_ord2_rgb_vs_rgb_dist}
correlate_pairwise_out2 <- function(data1, data2) {
  corrr::correlate(
    data1 %>% 
      rename(value1 = value) %>% 
      left_join(data2 %>% rename(value2 = value)) %>%
      select_if(is.numeric)
  )
}

correlate_pairwise_out2(
  colors_nbanfl_ord2_rgb_dist, 
  colors_nbanfl_ord2_rgb_tidy %>%
    do_pairwise_dist("manhattan") 
)
```


```{r colors_nbanfl_ord2_hsl_tidy}
colors_nbanfl_ord2_hsl_tidy <-
  colors_nbanfl_tidy_ord2 %>%
  add_hsl_cols() %>% 
  select(-hex) %>%
  tidyr::gather(rgb, value, h, s, l)
```

```{r colors_nbanfl_ord2_hsl_vs_hsl_dist}
correlate_pairwise_out2(
  colors_nbanfl_ord2_hsl_tidy %>%
    do_pairwise_dist("euclidean"),
  colors_nbanfl_ord2_hsl_tidy %>%
    do_pairwise_dist("manhattan")
)
```

```{r colors_nbanfl_ord2_rgb_vs_hsl_dist}
correlate_pairwise_out2(
  colors_nbanfl_ord2_rgb_tidy %>%
    do_pairwise_dist("euclidean"),
  colors_nbanfl_ord2_hsl_tidy %>%
    do_pairwise_dist("euclidean")
)
```

```{r colors_nbanfl_ord2_rgb_sim}
do_pairwise_similarity_byord <- function(data, ord) {
  data %>%
    filter(ord == ord) %>% 
    select(-ord) %>% 
    widyr::pairwise_similarity(name, rgb, value) %>% 
    rename(name1 = item1, name2 = item2) %>% 
    select(everything(), value = ncol(.)) %>% 
    mutate(ord = ord) %>% 
    select(ord, everything()) %>% 
    arrange(desc(value))
}

do_pairwise_similarity <- function(data, ords = c("primary", "secondary")) {
  bind_rows(
    data %>% 
      do_pairwise_similarity_byord(ords[1]),
    data %>% 
      do_pairwise_similarity_byord(ords[2])
  )
}

colors_nbanfl_ord2_rgb_sim <-
  colors_nbanfl_ord2_rgb_tidy %>% 
  do_pairwise_similarity()
```

```{r corrr_rgb_hsl_dist_sim}
cor_data <-
  colors_nbanfl_ord2_rgb_tidy %>% 
  do_pairwise_dist("euclidean") %>% 
  rename(rgb_dist = value) %>% 
  left_join(
    colors_nbanfl_ord2_hsl_tidy %>% 
      do_pairwise_dist("euclidean") %>% 
      rename(hsl_dist = value)
  ) %>%
  left_join(
    colors_nbanfl_ord2_rgb_tidy %>% 
      do_pairwise_similarity() %>% 
      rename(rgb_sim = value)
  ) %>% 
  left_join(
    colors_nbanfl_ord2_hsl_tidy %>% 
      do_pairwise_similarity() %>% 
      rename(hsl_sim = value)
  ) %>% 
  select_if(is.numeric)
corrr_rgb_hsl_dist_sim <- corrr::correlate(cor_data)

corrr_rgb_hsl_dist_sim %>% 
  corrr::network_plot(
    min_cor = 0, 
    colors = c("red", "green")
  )

GGally::ggcorr(cor_data, label = TRUE, label_round = 2)
```
[What is the best method for identifying similarity between colors?](https://stackoverflow.com/questions/5392061/algorithm-to-check-similarity-of-colors)

```{r colors_nbanfl_ord2_compare}
# colors_nbanfl_ord2_compare <- colors_nbanfl_ord2_rgb_dist
colors_nbanfl_ord2_compare <- colors_nbanfl_ord2_rgb_sim
# colors_nbanfl_ord2_compare <- 
#   colors_nbanfl_ord2_hsl_tidy %>% 
#   do_pairwise_similarity()
```

```{r colors_nbanfl_ord2_rgb_dist_debug, include = FALSE, echo = FALSE}
colors_nbanfl_ord2_compare %>% 
  group_by(ord, name1) %>% 
  do(slice(., c(1, n()))) %>% 
  ungroup()
colors_nbanfl_ord2_compare %>% 
  arrange(value) %>%
  filter(ord == "primary") %>% 
  filter(value != 0) %>% 
  slice(c(1:10, c(n()-9):n())) %>% 
  ungroup()
colors_nbanfl_ord2_rgb_tidy %>%
  filter(name %in% c("San Antonio Spurs", "Pittsburgh Steelers", "Brooklyn Nets")) %>%
  tidyr::spread(rgb, value)
```

Finally, I can visualize the distances using a "heat-map"-like plot.
In order to make the plot more insightful, I order the teams on each axis
by total distance from all other teams. (So teams with the most similarity to all
other teams appear towards the bottom and left, while teams that
contrast most with all others appear towards the top and right.)


```{r viz_colors_nbanfl_ord2_compare_byord_setup, include = FALSE}
add_nm_and_lg_cols <- function(data) {
  data %>%
  left_join(nms_nbanfl %>% select(name1 = name, league1 = league), by = c("name1")) %>%
  left_join(nms_nbanfl %>% select(name2 = name, league2 = league), by = c("name2"))
}

colors_nbanfl_ord2_compare_ranks <-
  colors_nbanfl_ord2_compare %>%
  add_nm_and_lg_cols() %>% 
  group_by(ord, name1) %>%
  summarize(value = sum(value)) %>%
  ungroup() %>%
  group_by(name1) %>% 
  summarize(value = sum(value)) %>% 
  ungroup() %>% 
  mutate(rank = row_number(value)) %>% 
  arrange(rank)
colors_nbanfl_ord2_compare_ranks
nms_nbanfl_ranks <-
  colors_nbanfl_ord2_compare_ranks %>% 
  left_join(
    nms_nbanfl %>% select(name1 = name, slug1 = slug, league),
    by = c("name1")
  )
lvls_nbanfl_ranks <-
  nms_nbanfl_ranks %>%
  pull(name1)

labs_nbanfl_ranks <-
  nms_nbanfl_ranks %>%
  pull(slug1)
```

```{r colors_nbanfl_ord2_compare_byord_viz, echo = FALSE}
create_colors_nbanfl_ord2_compare_viz <- function(data, ranks) {
  data %>% 
    # tidyr::complete(name1, name2) %>% 
    inner_join(
      ranks %>% select(name1, rank1 = rank),
      by = c("name1")
    ) %>% 
    inner_join(
      ranks %>% select(name2 = name1, rank2 = rank),
      by = c("name2")
    ) %>% 
    # mutate_at(vars(name1, name2), funs(factor(., levels = lvls_nbanfl_ranks)))
    mutate(name1 = reorder(name1, rank1),
           name2 = reorder(name2, rank2))
}
colors_nbanfl_ord2_compare_byord_viz <-
  colors_nbanfl_ord2_compare  %>%
  create_colors_nbanfl_ord2_compare_viz(colors_nbanfl_ord2_compare_ranks)
```


```{r colors_nbanfl_ord2_compare_byord_viz_debug, include = FALSE, eval = FALSE}
# colors_nbanfl_ord2_compare_viz %>% arrange(rank1)
# colors_nbanfl_ord2_compare_viz %>% arrange(name1)
# levels(colors_nbanfl_ord2_compare_viz$name1)
# levels(colors_nbanfl_ord2_compare_viz$name2)
```

```{r viz_colors_nbanfl_ord2_compare_byord, echo = FALSE}
lab_subtitle_nbanfl <-
  stringr::str_replace_all(lab_subtitle_kclusts, lab_lgs, "NBA and NFL")
lab_caption_nbanfl <-
  paste0(
    "Dark = similar; light = dissimilar.",
    "Similarity is based on Euclidean distance of rgb values.\n",
    " Order is determined by overall similarity across both primary and secondary colors.",
    "\n Top 2% similar and dissimilar team-team colors emphasized with black and white respectively."
  )
viz_labs_colors_nbanfl_ord2_compare <-
  labs(
    title = "Pairwise Distance of Colors",
    subtitle = lab_subtitle_nbanfl,
    caption = lab_caption_nbanfl,
    x = NULL,
    y = NULL
  )

visualize_colors_nbanfl_ord2_compare_const <-
  function(data) {
    data_pranked <-
      data %>% 
      mutate(value_prank = percent_rank(value))
    data %>% 
      # filter(league1 == league2) %>%
      # mutate(lg_ord = paste0(toupper(league1), ", ", stringr::str_to_title(ord))) %>% 
      mutate(value = ifelse(is.na(value), 0, value)) %>% 
      ggplot(aes(x = name1, y = name2)) +
      geom_tile(aes(fill = value)) +
      scale_x_discrete(labels = labs_nbanfl_ranks) +
      scale_y_discrete(labels = labs_nbanfl_ranks) +
      viridis::scale_fill_viridis(option = "E") +
      # scale_fill_gradient(low = "black", high = "white") +
      geom_tile(
        data = data_pranked %>%
          filter(value_prank <= 0.02),
        fill = "green"
      ) +
      geom_tile(
        data = data_pranked %>%
          filter(value_prank >= 0.98),
        fill = "red"
      ) +
      viz_theme_base +
      theme(axis.text.x = element_text(angle = 90)) +
      viz_labs_colors_nbanfl_ord2_compare
  }

viz_colors_nbanfl_ord2_compare_byord <-
  colors_nbanfl_ord2_compare_byord_viz %>%
  visualize_colors_nbanfl_ord2_compare_const() +
  facet_wrap(~ ord, scales = "free")
viz_colors_nbanfl_ord2_compare_byord
```


```{r viz_colors_nbanfl_ord2_compare, echo = FALSE}
colors_nbanfl_ord2_compare_viz <-
  colors_nbanfl_ord2_compare %>%
  group_by(name1, name2, league1, league2) %>% 
  summarize(value = sum(value)) %>% 
  ungroup() %>% 
  create_colors_nbanfl_ord2_compare_viz(colors_nbanfl_ord2_compare_ranks)

viz_colors_nbanfl_ord2_compare <-
  colors_nbanfl_ord2_compare_viz %>%
  visualize_colors_nbanfl_ord2_compare_const()
viz_colors_nbanfl_ord2_compare
```

